Distrivia: A Distributed Trivia Game

### 2011-05-06 papers/phase-three/paper.tex

%\begin{abstract}
%In this paper we detail our current progress on our distributed trivia game.
%We will explain what has been done so far, what we need to do, and how we
%have met the requirements for the project. We also discuss the architecture
%for our system.
%\end{abstract}

\section{Project Motivation}
\section{Architecture}
\begin{figure}[h!]
  \centering
    \includegraphics[scale=0.4]{diagram.png}
   \caption{High level overview of the system}
 \end{figure}

\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host running Monit\cite{monit}. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage whether or not the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.

\subsection{Server Applications}
The server is currently almost feature complete, but still needs a lot of
testing. Client registration, and login are all supported. A client can
join a game, retrieve questions for that game, and answer questions. Both a
global leader board and individual per game leader board can also be requested.
We recently refactored our API to be more universally consistent.
The servers and clients now work on a more consistent basis where the clients all expect the same information and use the same calls.

>>>>

\begin{abstract}
In this paper we will discuss our decisions and detail our discoveries throughout designing a distributed trivia system.
\end{abstract}

\section{Project Motivation}
Our motivation for this project was to build a system that would allow people to compete across multiple platforms reliably.
High availability was a major concern for us, as it enables people to compete at all times.
This form of trivia is a multiplayer social game.
It takes advantage of a number of different platforms so that users can feel free to play on any device they prefer.
The web client ensures that almost any device will be able to connect and play.
Designing it on mobile platforms allows us to take advtange of the specific platform and lets people easily play on the go.
\section{Architecture}
\subsection{Servers}
For our servers, we are using Amazon EC2 micro instances.
These are running x86-64 Amazon Linux.
They have 8gb of network attached storage and 613mb of memory.

Each server is installed with our custom web app that services our requests.
Each one is also equipped with an instance of Riak for maintaining our database.
The Python web app utilizes the Flask web framework and we serve the application and all files with the Tornado web server, both of which are open source.
We have five servers all running our server application and hosting the databases.
Two of those are set up as hot spares in case of a failure.

Currently, we are using the NGINX \cite{nginx} in a reverse proxy configuration for our load sharing needs.
For this project there are two load balancers which are running on personal desktop computers that have reliable internet connections.
These are set up with round robin DNS entries, courtesy of ZoneEdit.com, a free DNS management service.
\section{Design}
\begin{figure}[h!]
  \centering
    \includegraphics[scale=0.4]{diagram.png}
   \caption{High level overview of the system}
 \end{figure}
\subsection{Servers}
For this project we deployed six servers running on Amazon's Elastic Compute Cloud \cite{aec}.
One server is being used as a web-monitor host running Monit\cite{monit}.
We were able to login to the web-monitor to see the status of all our servers, databases, and load sharers.
This assisted us in managing whether or not the services were running and allowed us to track down any problems, if they arose.
Each server contains the server application that responds to clients as well as the Riak \cite{riak} databases.
Riak handles auto-synchronization of the databases between the five servers.
Note that the hot spares were setup to be part of the database cluster so they have a full copy of the database on hand, however they did not accept any reads or writes directly.
This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update Riak and it would sync the databases to any additional servers as well.
These servers, therefore, all contain the same information.
If any servers go down, the distributed system would still run fine.
We could have all servers go down except one and still be reliably handling requests.

\subsection{Server Applications}
The server application handles requests from the user via the load sharers.
The server application supports the following features:
\begin{itemized}
\item Client registration
\item Login
\item Query games status
\item Joining public games
\item Joining private games
\item Creating private games
\item Question retrieval
\item Question answering
\item Global leaderboard
\end{itemized}
